Advances in Artificial Intelligence for Data Visualization: Developing Computer Vision Models to Automate Reading of Data Plots, with Application to Predictive Model Diagnostics

### 2022-03-09 Milestone_1/01-chap1.Rmd
add overview of the thesis

# Structure of the Thesis (need disscussion)

An overview of the thesis, including some brief detail of the proposed content of all chapters. This should encompass a description of the theoretical and conceptual framework that underlies the thesis, and the procedures that are to be used in addressing the research questions

The

The first project is

1. Automatic Visual Statistical Inference, with Application to Linear Regression Diagnostics
2. ...
3. ...

# Automatic Visual Statistical Inference, with Application to Linear Regression Diagnostics

@@

>>>>

# Research Problem

<!-- 1. Provide visual diagnostics for projection pursuit optimisation to examine whether the optimisers are able to find the optima, compare optimisation paths, inspect optimisers’ coverage of the space, and detect anomalies in the algorithms. -->

<!-- 2. Extend and enhance the visual diagnostic tools developed in the first project to stochastic optimal control problems, which is a multistage optimisation problem with more than two exogenous variables. -->

<!-- 3. Improve the interpretability and explainability of stochastic optimal control algo- rithms by identifying the most important drivers in the optimal decisions suggested by the optimisers and studying how the decision changes across time. -->

The main objective of this research is to build an automatic visual inference system for the purpose of conducting large-scale visual test. The research will focus on three projects as follows:

1. Propose a prototype of automatic visual inference system for evaluating lineups of residual plots of the classical normal linear regression model and study factors that affect the performance of the system with comparison to human subjects.

2. Extend and enhance the automatic visual inference system proposed in the first project to general lineup protocol problems by adopting image similarity assessed by computer vision models.

3. Validate and apply the automatic visual inference system to predictive model diagnostics on empirical data.

# Overview of the Thesis

The first project proposes a prototype of automatic visual inference system for evaluating lineups of residual plots, with a focus on the classical normal linear regression model given it is one of the simplest predictive models. The automatic system is identical to the lineup protocol except evaluators are replaced by computer vision models. The computer vision models are trained by using data simulated from linear models with violations of different classical assumptions, such as linearity and homoscedasticity. Data of human performance on evaluating residual plots generated under the same simulation setting is collected by conducting online human subject experiments, such that the comparison between the power of human subject-assessed visual tests and the system-assessed visual tests can be made. Finally, factors that affect the performance of the automatic system are studied for improving the architecture and design of the computer vision models.

The second project aims to extend the proposed automatic system to lineup protocol with no assumptions on the type of the data plot and the data generating process. It only requires the protocol to asks the evaluators to select the most different plot(s) from the lineup. The automatic system will be built upon image similarity produced by computer vision models.

The third project aims to apply the automatic system to predictive model diagnostics on empirical data. Potential predictive models considered by this research are linear model, generalized linear models and non-parametric regression. Multiple human subject experiments will need to be conducted to validate the effectiveness of the automatic system.

# Automatic Visual Statistical Inference for linear regression diagnostics

@@

# General Purpose Automatic Visual Statistical Inference System

# Visual Diagnostics for Predictive Models with ...

### 2022-03-09 Milestone_1/_book/thesis.tex
add overview of the thesis

\hypertarget{structure-of-the-thesis-need-disscussion}{%
\chapter{Structure of the Thesis (need disscussion)}\label{structure-of-the-thesis-need-disscussion}}

An overview of the thesis, including some brief detail of the proposed content of all chapters. This should encompass a description of the theoretical and conceptual framework that underlies the thesis, and the procedures that are to be used in addressing the research questions

The

The first project is

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Automatic Visual Statistical Inference, with Application to Linear Regression Diagnostics
\item
  \ldots{}
\item
  \ldots{}
\end{enumerate}

\hypertarget{automatic-visual-statistical-inference-with-application-to-linear-regression-diagnostics}{%
\chapter{Automatic Visual Statistical Inference, with Application to Linear Regression Diagnostics}\label{automatic-visual-statistical-inference-with-application-to-linear-regression-diagnostics}}

@@

>>>>

\hypertarget{research-problem}{%
\chapter{Research Problem}\label{research-problem}}

The main objective of this research is to build an automatic visual inference system for the purpose of conducting large-scale visual test. The research will focus on three projects as follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Propose a prototype of automatic visual inference system for evaluating lineups of residual plots of the classical normal linear regression model and study factors that affect the performance of the system with comparison to human subjects.
\item
  Extend and enhance the automatic visual inference system proposed in the first project to general lineup protocol problems by adopting image similarity assessed by computer vision models.
\item
  Validate and apply the automatic visual inference system to predictive model diagnostics on empirical data.
\end{enumerate}

\hypertarget{overview-of-the-thesis}{%
\chapter{Overview of the Thesis}\label{overview-of-the-thesis}}

The first project proposes a prototype of automatic visual inference system for evaluating lineups of residual plots, with a focus on the classical normal linear regression model given it is one of the simplest predictive models. The automatic system is identical to the lineup protocol except evaluators are replaced by computer vision models. The computer vision models are trained by using data simulated from linear models with violations of different classical assumptions, such as linearity and homoscedasticity. Data of human performance on evaluating residual plots generated under the same simulation setting is collected by conducting online human subject experiments, such that the comparison between the power of human subject-assessed visual tests and the system-assessed visual tests can be made. Finally, factors that affect the performance of the automatic system are studied for improving the architecture and design of the computer vision models.

The second project aims to extend the proposed automatic system to lineup protocol with no assumptions on the type of the data plot and the data generating process. It only requires the protocol to asks the evaluators to select the most different plot(s) from the lineup. The automatic system will be built upon image similarity produced by computer vision models.

The third project aims to apply the automatic system to predictive model diagnostics on empirical data. Potential predictive models considered by this research are linear model, generalized linear models and non-parametric regression. Multiple human subject experiments will need to be conducted to validate the effectiveness of the automatic system.

\hypertarget{automatic-visual-statistical-inference-for-linear-regression-diagnostics}{%
\chapter{Automatic Visual Statistical Inference for linear regression diagnostics}\label{automatic-visual-statistical-inference-for-linear-regression-diagnostics}}

@@

\hypertarget{general-purpose-automatic-visual-statistical-inference-system}{%
\chapter{General Purpose Automatic Visual Statistical Inference System}\label{general-purpose-automatic-visual-statistical-inference-system}}

\hypertarget{visual-diagnostics-for-predictive-models-with}{%
\chapter{Visual Diagnostics for Predictive Models with \ldots{}}\label{visual-diagnostics-for-predictive-models-with}}
